---
title: "hazard-model-denis"
author: "Denis Bykov"
date: "March 6, 2018"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(survival)
library(survminer)

# Import RDS File
temporal = readRDS("Hazard_Temporal_Data.rds")
temporal = subset(temporal, select = -c(Year, Date, Month_event, X, ApprovalDate, BorrZip, ProjectCounty,ProjectState))
```

## Univariate Hazard Model

Understanding results of a single covariate on survival. Chargeoff is modeled as 1, so positive coefficients means more correlated with having a chargeoff.

```{r univariate hazard model, echo=FALSE}
# Univariate Hazard Model
res_one.cox <- coxph(Surv(TimeElapsed, result) ~ zill_ZHVI, data = temporal)
# res.cox
summary(res_one.cox)
```

## Univariate Hazard Applied to Multiple Covariates
Alternative to running multivariate model, studies impact of set of variables but does each one individually. Used for looking at impact of covariates that produce NAs for some reason in multivariate model.

```{r multi-univariate, echo=FALSE}
covariates <- c("ed_st_LOCAL_REVENUE", "ed_cty_TLOCREV", "zill_SizeRank_med", "zill_SizeRank_inc", "zill_MedianValue", "sp_Adj_Close","ag_Total")
univ_formulas <- sapply(covariates,
                        function(x) as.formula(paste('Surv(TimeElapsed, result)~', x)))
univ_models <- lapply( univ_formulas, function(x){coxph(x, data = temporal)})
# Extract data 
univ_results <- lapply(univ_models,
                       function(x){ 
                         x <- summary(x)
                         p.value<-signif(x$wald["pvalue"], digits=2)
                         wald.test<-signif(x$wald["test"], digits=2)
                         beta<-signif(x$coef[1], digits=2);#coeficient beta
                         HR <-signif(x$coef[2], digits=2);#exp(beta)
                         HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                         HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                         HR <- paste0(HR, " (", 
                                      HR.confint.lower, "-", HR.confint.upper, ")")
                         res<-c(beta, HR, wald.test, p.value)
                         names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                       "p.value")
                         return(res)
                         #return(exp(cbind(coef(x),confint(x))))
                       })
res <- t(as.data.frame(univ_results, check.names = FALSE))
knitr::kable(res)
# as.data.frame(res)
```

## Multivariate Hazard Model

Cox model across all covariates except those that produced NAs results in previous runs. Looks at effect of each variable in aggregate versus individually as done above.

```{r multivariate, echo=FALSE}
# omit = paste(na_var, "Year", "Date", "Month_event", "X", "ApprovalDate", "BorrZip", "ProjectCounty", "ProjectState", sep = " - ")
# na_var = paste("ed_st_LOCAL_REVENUE", "ed_cty_TLOCREV", "zill_SizeRank_med", "zill_SizeRank_inc", "zill_MedianValue", "unemp_rate_US","BCI","CCI","sp_Adj_Close","sp_Returns","sp_Volume","sp_Close","sp_Low","sp_High","sp_Open","ag_Total","ag_Private_dollar","ag_Public_dollar","ag_Private_nom","ag_Public_nom", sep = " - ")
res.cox <- coxph(Surv(TimeElapsed, result) ~ . , data =  temporal)
summary(res.cox)
```

```{r, echo=FALSE}
# Plot the baseline survival function
ggsurvplot(survfit(res.cox), palette= '#2E9FDF',
           ggtheme = theme_minimal(), data = temporal)
```

```{r, echo=FALSE}
# Create the new data  
res2.cox <- coxph(Surv(TimeElapsed, result) ~ unemp_rate_cty + CCI + zill_ZHVI , data =  temporal)

# unemp_rate_cty, zill_zhvi, cci
unemp_vals = unique(temporal$unemp_rate_cty)[-1]
len = length(unemp_vals)
df <- with(temporal, data.frame(unemp_rate_cty = c(0.5,40.5), CCI = rep(mean(CCI, na.rm = TRUE), 2), zill_ZHVI = rep(mean(zill_ZHVI, na.rm = TRUE), 2)))

# Survival curves
fit <- survfit(res2.cox, newdata = df)
ggsurvplot(fit, conf.int = TRUE, legend.labs=c("Unemp=0.5", "Unemp=40.5"), ggtheme = theme_minimal(), data = temporal)

```

## Checking the Significance of Income
Income data contains many missing values, wanted to check using multi-univariate in case valuable to include that data.

```{r, echo=FALSE}
# Load Income data
temporal2 = readRDS("Hazard_Temporal_Income.rds")
temporal2 = subset(temporal2, select = colnames(temporal2)[c(45:66)])
```


```{r, echo=FALSE}
# Working with Temporal2
income_vars = c("income_n1_2","income_prep_2","income_n2_2","income_a00100_2","income_a00200_2","income_a00300_2","income_a00600_2","income_a00900_2","income_a18300_2","income_a19700_2","income_n1_3","income_prep_3","income_n2_3","income_a00100_3","income_a00200_3","income_a00300_3","income_a00600_3","income_a00900_3","income_a18300_3","income_a19700_3")

univ_formulas <- sapply(income_vars,
                        function(x) as.formula(paste('Surv(TimeElapsed, result)~', x)))
univ_models <- lapply( univ_formulas, function(x){coxph(x, data = temporal2)})
# Extract data 
univ_results <- lapply(univ_models,
                       function(x){ 
                         x <- summary(x)
                         p.value<-signif(x$wald["pvalue"], digits=2)
                         wald.test<-signif(x$wald["test"], digits=2)
                         beta<-signif(x$coef[1], digits=2);#coeficient beta
                         HR <-signif(x$coef[2], digits=2);#exp(beta)
                         HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                         HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                         HR <- paste0(HR, " (", 
                                      HR.confint.lower, "-", HR.confint.upper, ")")
                         res<-c(beta, HR, wald.test, p.value)
                         names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                       "p.value")
                         return(res)
                         #return(exp(cbind(coef(x),confint(x))))
                       })
res_income <- t(as.data.frame(univ_results, check.names = FALSE))
knitr::kable(res_income)

```

## Multivariate Analysis of Income Data

```{r, echo=FALSE}
res_income.cox <- coxph(Surv(TimeElapsed, result) ~ . , data =  temporal2)
summary(res_income.cox)

```

## Graphs with Income Data
Taking the most statistically significant groups and creating plots.
```{r, echo=FALSE}
# Create the new data
# income_a00900_3 income_a00300_2 income_a00300_3 income_prep_3 income_n1_2
res_income2.cox <- coxph(Surv(TimeElapsed, result) ~ income_a00900_3 + income_a00300_2 + income_a00300_3 + income_prep_3 + income_n1_2, data =  temporal2)

df_income <- with(temporal2, 
  data.frame(
    income_a00900_3 = c(-10850330,30398548),
    income_a00300_2 = rep(mean(income_a00300_2, na.rm = TRUE), 2),
    income_a00300_3 = rep(mean(income_a00300_3, na.rm = TRUE), 2),
    income_prep_3 = rep(mean(income_prep_3, na.rm = TRUE), 2),
    income_n1_2 = rep(mean(income_n1_2, na.rm = TRUE), 2) 
  )
)

# Survival curves
fit_income <- survfit(res_income2.cox, newdata = df_income)
ggsurvplot(fit_income, conf.int = TRUE, legend.labs=c("a00900=-10850330", "a00900=30398548"), ggtheme = theme_minimal(), data = temporal2)


```

```{r, echo=FALSE}
res_income2.cox <- coxph(Surv(TimeElapsed, result) ~ income_a00900_3 + income_a00300_2 + income_a00300_3 + income_prep_3 + income_n1_2, data =  temporal2)

df_income <- with(temporal2, 
  data.frame(
    income_a00900_3 = rep(mean(income_a00900_3, na.rm = TRUE), 2),
    income_a00300_2 = c(min(income_a00300_2, na.rm = TRUE), max(income_a00300_2, na.rm = TRUE)),
    income_a00300_3 = rep(mean(income_a00300_3, na.rm = TRUE), 2),
    income_prep_3 = rep(mean(income_prep_3, na.rm = TRUE), 2),
    income_n1_2 = rep(mean(income_n1_2, na.rm = TRUE), 2) 
  )
)

# Survival curves
fit_income <- survfit(res_income2.cox, newdata = df_income)
ggsurvplot(fit_income, conf.int = FALSE, legend.labs=c("a00300_2=Min", "a00300_2=Max"), ggtheme = theme_minimal(), data = temporal2)
```


```{r, echo=FALSE}
res_income2.cox <- coxph(Surv(TimeElapsed, result) ~ income_a00900_3 + income_a00300_2 + income_a00300_3 + income_prep_3 + income_n1_2 + income_a00600_2 + income_a00100_2 + income_prep_2, data =  temporal2)

df_income <- with(temporal2, 
  data.frame(
    income_a00900_3 = rep(mean(income_a00900_3, na.rm = TRUE), 2),
    income_a00300_2 = rep(mean(income_a00300_2, na.rm = TRUE), 2),
    income_a00300_3 = rep(mean(income_a00300_3, na.rm = TRUE), 2),
    income_prep_3 = rep(mean(income_prep_3, na.rm = TRUE), 2),
    income_n1_2 = rep(mean(income_n1_2, na.rm = TRUE), 2),
    income_a00600_2 = rep(mean(income_a00600_2, na.rm = TRUE), 2),
    income_a00100_2 = c(min(income_a00100_2, na.rm = TRUE), max(income_a00100_2, na.rm = TRUE)),
    # income_a00100_2 = rep(mean(income_a00100_2, na.rm = TRUE), 2),
    income_prep_2 = rep(mean(income_prep_2, na.rm = TRUE), 2)
  )
)

# Survival curves
fit_income <- survfit(res_income2.cox, newdata = df_income)
ggsurvplot(fit_income, conf.int = FALSE, legend.labs=c("a00100_2=Min", "a00100_2=Max"), ggtheme = theme_minimal(), data = temporal2)
```

```{r, echo=FALSE}
res_income2.cox <- coxph(Surv(TimeElapsed, result) ~ income_a00900_3 + income_a00300_2 + income_a00300_3 + income_prep_3 + income_n1_2 + income_a00600_2 + income_a00100_2 + income_prep_2, data =  temporal2)

a = sort(unique(temporal2$income_a00100_2)[-1])

df_income <- with(temporal2, 
  data.frame(
    income_a00900_3 = rep(mean(income_a00900_3, na.rm = TRUE), 3),
    income_a00300_2 = rep(mean(income_a00300_2, na.rm = TRUE), 3),
    income_a00300_3 = rep(mean(income_a00300_3, na.rm = TRUE), 3),
    income_prep_3 = rep(mean(income_prep_3, na.rm = TRUE), 3),
    income_n1_2 = rep(mean(income_n1_2, na.rm = TRUE), 3),
    income_a00600_2 = rep(mean(income_a00600_2, na.rm = TRUE), 3),
    # income_a00100_2 = c(a[1], a[150], a[300]),
    income_a00100_2 = c( quantile(a, c(0, .25, .5))[[1]], quantile(a, c(0, .25, .5))[[2]], quantile(a, c(0, .25, .5))[[3]]  ),
    income_prep_2 = rep(mean(income_prep_2, na.rm = TRUE), 3)
  )
)

# Survival curves
fit_income <- survfit(res_income2.cox, newdata = df_income)
ggsurvplot(fit_income, conf.int = FALSE, legend.labs=c("a00100_2=Min", "a00100_2=25%", "a00100_2=Median"), ggtheme = theme_minimal(), data = temporal2)
```


https://www.kaggle.com/irs/individual-income-tax-statistics/data 

## To Do
Cleanup NAs, make sure same formatting. Normalize. Take log of big numbers?